---
title: "Explore_transect_community"
author: "Aimara Planillo"
date: "3/30/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Summary and Plots of transect diversity data

### packages
```{r}
library(sf)
library(ggplot2)
library(tidyr)
library(dplyr)
library(FD)
library(BiodiversityR)
library(gridExtra) # to do the final arrangement of the plots
```

### workspace
```{r}
WorkDir <- getwd()

Data_wd <- file.path(WorkDir, "2_Raw_data/")
Plots_wd <- file.path(WorkDir, "5_Results/Plots_birds")
```

# Get data summaries transect

```{r load data for summaries}
# load transects and get length
transects <- read_sf(paste0(Data_wd,
                            "/geodata/birds_transects_geometry_25833.gpkg"))
transects
plot(st_geometry(transects))
transects.lengths <- st_length(transects)/1000
transects.lengths2 <- cbind.data.frame(site = transects$routcode_2, 
                                    length = as.numeric(transects.lengths))
# order transects as
transects.lengths2 <- transects.lengths2[order(transects.lengths2$site),]

# load bird data
birds.raw <- read.csv(paste0(Data_wd, "/Birds_sitexsp_fullnames.csv"))
head(birds.raw)

birds.abund <- rowSums(birds.raw[-1])
birds.sp <- rowSums(birds.raw[-1] > 0)

birds.summary <- cbind.data.frame(site = tolower(birds.raw$SITE), 
                                  bird.sp = birds.sp,
                                  bird.ind = birds.abund)


# load env data
env100m.tmp <- read.csv("./2_Raw_data/transects_allenvir_100m.csv")

# renaming covariates for the analysis
abiotic.covariates <- cbind.data.frame(site = env100m.tmp$routecode_2, 
                                       tree = env100m.tmp$tree_cover,
                                       o.green = env100m.tmp$open_green,
                                       noise = env100m.tmp$noise,
                                       pop = env100m.tmp$pop_100m)
# prey variables
carabids <- read.csv("./2_Raw_data/carabids_summary_pois_predictions.csv")
grasshoppers <- read.csv("./2_Raw_data/grasshoppers_summary_pois_predictions.csv")
spiders <- read.csv("./2_Raw_data/spiders_summary_pois_predictions.csv")

prey.abu <- cbind.data.frame(site = carabids$site,
                             prey = carabids$Abu + grasshoppers$Abu + spiders$Abu)

transects.variables <- left_join(transects.lengths2, birds.summary, by = "site") %>% 
  mutate("ind/km" = bird.ind/length) %>% 
  left_join(abiotic.covariates, by = "site") %>% 
  left_join(prey.abu, by = "site")

transects.variables
write.csv(transects.variables, "./4_Processed_data/Transects_variables.csv")
```


# Get covariates summaries by transect to plot

```{r modify data format}
toplot.1 <- transects.variables %>% 
  mutate(S = bird.sp, A = bird.ind, T = tree, OG = o.green, 
         N = noise, HP = pop) %>%
  gather(my.variable, label.value, S:HP) %>%
  dplyr::select(c(site, my.variable, label.value))

my.values <- transects.variables %>% 
  mutate(S = bird.sp/max(bird.sp), A = bird.ind/max(bird.ind), T = tree/max(tree), 
         OG = o.green/max(o.green), N = noise/max(noise), HP = pop/max(pop)) %>%
  gather(my.variable, my.value, S:HP)  %>%
  dplyr::select(site, my.variable, my.value)


my.values
toplot <- cbind.data.frame(toplot.1, my.value = my.values$my.value)

toplot$my.variable <- factor(toplot$my.variable, levels = c("S", "A", "T", "OG", "N", "HP"))
```

```{r plot variables}
colnames(toplot)

# subset for testing
species.variables <- c('S')

test <- toplot %>%
  filter(my.variable %in% species.variables) %>%
  filter(site == "be2") 

ggplot(data=test, aes(x=0, y = my.value))+
  geom_bar(stat="identity") 

ggplot(data=test, aes(x=0, y = my.value)) +
  geom_bar(stat="identity", width = 0.8, fill = "blue") +
  geom_text(aes(label = round(my.value,1), x = 0.5, y = 0.1), 
            size = 30, vjust = 0, hjust = 0) +
  geom_text(aes(label = as.character(my.variable)), x = -0.5, y = 0.1, 
            size = 30, vjust = 0, hjust = 1) +
  xlim(c(-2,2)) +
  ylim(c(0,1)) +
  theme_void() # This widens the right margin

# for each transect
transects.names <- unique(toplot$site)  
myvariable.names <- c("S","A","T", "OG", "N", "HP")

test <- toplot %>%
  filter (site == transects.names[1]) 

ggplot(data = test, aes(x=0, y = my.value)) + 
      geom_bar(stat="identity",  width = 0.5, aes(fill = my.variable)) +
    scale_fill_manual(values = c("darkblue", "lightblue", "darkgreen", "lightgreen", "grey", "darkorange"))+
  geom_text(aes(label = round(label.value,1), x = 0.4, y = 0), 
            size = 15, vjust = 0, hjust = 0) +
  geom_text(aes(label = as.character(my.variable)), x = -0.4, y = 0, 
            size = 15, vjust = 0, hjust = 1) +
  facet_wrap( ~ my.variable, ncol=2) +
  xlim(c(-2,2))+
  ylim(c(0,1)) +
  theme_void() +
  theme(strip.text.x = element_blank(), 
        legend.position = "none")
  
# in a loop for all the transects
my.variables1 <- c("S", "A")
my.variables2 <- c("T", "OG", "N", "HP")

for(i in 1:length(transects.names)){
  # subset of the transect
  test <- subset(toplot, toplot$site == transects.names[i])
  # create upper plot: S, A
  test1 <- test %>%
     filter(my.variable %in% my.variables1)
 
  plot1 <- ggplot(data = test1, aes(x=0, y = my.value)) + 
      geom_bar(stat="identity",  width = 0.5, aes(fill = my.variable)) +
    scale_fill_manual(values = c("darkblue", "lightblue")) +
    geom_text(aes(label = round(label.value,1), x = 0.4, y = 0), 
            size = 12, vjust = 0, hjust = 0, color = "black") +
    geom_text(aes(label = as.character(my.variable)), x = -0.4, y = 0, 
            size = 14, vjust = 0, hjust = 1, fontface = "bold") +
    facet_wrap( ~ my.variable, ncol=2) +
    xlim(c(-2,2))+
    ylim(c(0,1)) +
    theme_void() +
    theme(strip.text.x = element_blank(), 
        legend.position = "none")

  # create lower plot: T, OG, N, HP
  test2 <- test %>%
     filter(my.variable %in% my.variables2)
 
  plot2 <- ggplot(data = test2, aes(x=0, y = my.value)) + 
        geom_bar(stat="identity",  width = 0.5, aes(fill = my.variable)) +
        scale_fill_manual(values = c("darkgreen", "limegreen","grey30", "tomato4")) +
        geom_text(aes(label = round(label.value,1), x = 0.4, y = 0), 
                size = 10, vjust = 0, hjust = 0, color = "grey40") +
        geom_text(aes(label = as.character(my.variable), color = my.variable), x = -0.4, y = 0, 
                size = 12, vjust = 0, hjust = 1) +
        scale_color_manual(values = c("darkgreen", "limegreen","grey30", "tomato4"))+
        facet_wrap( ~ my.variable, ncol=2) +
        xlim(c(-2,2))+
        ylim(c(0,1)) +
        theme_void() +
        theme(strip.text.x = element_blank(), 
              legend.position = "none")

  # create white space
  blank <- rectGrob(gp=gpar(col="white")) # make a white spacer grob

  # put everything togheter and save
  plot.full <- grid.arrange(plot1, blank, plot2, heights=c(0.5, 0.2, 1), nrow=3)  
  ggsave(paste0(Plots_wd, "/", transects.names[i], "_variables_summary_lowres.png"), 
       width = 6, height = 3, dpi = 300,
       plot = plot.full)
}
```


# Get diversity by transect: RAC 

Rank abundance curves (RAC): A rank abundance curve or Whittaker plot displays relative species abundance. It can also be used to visualize species richness and species evenness. The curve is a 2D chart with relative abundance on the Y-axis and the abundance rank on the X-axis.

```{r load data}
# birds
birds <- read.csv("./4_Processed_data/NEWBirds_abund_noexo_norare2_noaquatic_bykm.csv")
birds$site <- as.factor(tolower(birds$site))
str(birds)
head(birds)
ncol(birds)

# we put site as row name instead of a column
row.names(birds) <- birds$site
birds.abund <- birds %>% 
  select(-site)
birds.abund
```

```{r get rac}
# rank abundance for one site
levels(birds$site)[1]
# [1] "be10"

rankabundance(x = birds.abund,  
              y = birds, 
              factor = "site", 
              level = levels(birds$site)[1])

# Do all the transects
transects <- levels(birds$site)

rankabund.list <- list()
for (i in 1:length(transects)){
  rankabund.list[[i]] <- cbind.data.frame(rankabundance(x = birds.abund, 
                                                        y = birds, 
                                                        factor = "site", 
                                                        level = transects[i]), 
                                          transect = transects[i])
}
names(rankabund.list) <- transects
rankabund.list$be10
rankabund.table <- do.call("rbind", rankabund.list)
```

```{r plot rac}
# default plotting
rankabunplot(rankabund.list$be10, scale = 'logabun', addit = FALSE, specnames = FALSE, 
             col = "red", lwd = 2, ylim = c(0.5,100))


# Plotting manually

# ggplot all curves
ggplot(data = rankabund.table, aes(x = rank, y = abundance, color = transect)) +
  geom_line()

# all plots together
rank.curves <- ggplot(data = rankabund.table, aes(x = rank, y = abundance)) +
  geom_line(col = "grey20", lwd = 2) +
  facet_wrap(~transect, ncol = 5) + #, scales = "free_y") +
  theme_bw() +
  theme(panel.grid = element_blank(), 
        axis.text = element_text(size = 14), 
        axis.title = element_text(size = 16), 
        strip.text.x = element_text(size= 14, face = "bold"))
rank.curves

# plots 1 by 1. Saved as a list
plot_list = list()
for (i in 1:length(transects)){
my.data <- subset(rankabund.table, transect  == transects[i])
p <- ggplot(data = my.data, aes(x = rank, y = abundance)) +
    geom_line(col = "grey20", lwd = 2) +
    scale_y_continuous("abundance", breaks = c(0,60), limits=c(0,61)) +
    scale_x_continuous(breaks = c(0, 50), labels = c(0, 50), limits = c(0,51)) +
    # ylim(c(0,60)) +
    # xlim(c(0,45)) +
    ggtitle("RAC") +
    theme(panel.grid = element_blank(), 
          panel.background = element_blank(),
          axis.text = element_text(size = 40), 
          axis.title = element_text(size = 40),
          # axis.title = element_blank(),
          axis.line = element_line("grey30", size = 1.5), 
          plot.title = element_text(size = 50, hjust = 0.5 ))
  print(p)
  ggsave(paste0(Plots_wd, "/rac_plot_", transects[i], ".png"), 
         width = 6, height = 6, dpi = 300)
}
```


# Get transects functional diversity 
community-level weighted means of trait values

```{r get community weighted means of traits}
traits.tmp <- read.csv("./4_Processed_data/NEWBirds_traits_noexo_norare2_noaquatic.csv")
head(traits)
head(birds)

# select the used traits
traits.birds <- traits.tmp %>%
  mutate(log10.bodymass = log(bodymass.value,10)) %>% 
  select(diet.5cat, migrant, log10.bodymass)

rownames(traits.birds) <- traits.tmp$scientific2


cwm.birds <- functcomp(as.matrix(traits.birds), as.matrix(birds.abund), CWM.type = "dom")

write.csv(cwm.birds, "./4_Processed_data/Transects_community_weighted_means_traits.csv")

```


# Plot transects in the study area 

### Packages
```{r}
library(raster)
library(sf)
library(tmap)
library(RColorBrewer)
```

```{r load gis data}
# berlin outside
berlin.out <- read_sf(paste0(Data_wd, "/geodata/Berlin_border_3035.gpkg"), 
                      crs = 3035)
berlin.out <- st_transform(berlin.out, crs = 25833)

grid.birds <- read_sf(paste0(Data_wd, "/geodata/birds_quadrants1x1km_3035.gpkg"), crs = 3035)
grid.birds <- st_transform(grid.birds, crs = 25833)


plot(st_geometry(berlin.out), main = "Berlin")
plot(st_geometry(grid.birds), add = TRUE, col = "blue")

# imperviousness rasters
imperv.tmp <- raster(paste0(Data_wd, "/geodata/impervious_surface_20m_25833.tif"), 
                     crs = 25833)
```

```{r plot study area map}
imperv.crop <- mask(imperv.tmp, berlin.out)

my.palette <- c("palegoldenrod", "sandybrown", "#660000")

tmap_mode("plot")

(imperv.map <- tm_shape(imperv.crop) +
       tm_raster(palette = my.palette, title = "Impervious surface (%)") +

  tm_shape(berlin.out) +
  tm_borders("black", lwd = 2) +
  tm_shape(grid.birds) +
  tm_borders("white", lwd = 5) +
  tm_shape(grid.birds) +
  tm_borders("black", lwd = 2) +
  tm_text("routcode_2", col = "black", size = 1,
          remove.overlap = FALSE, 
          bg.color = "palegoldenrod", bg.alpha = 0.5,
          fontface = "bold",
          just = "left", xmod = +0.6, ymod = +0.2) +
  tm_layout(main.title = "study area", 
            legend.position = c(0.75, 0.7),
            legend.bg.color = "white", 
            legend.frame = "black",
            legend.text.size = 0.6,
            legend.title.size = 1,
            frame = FALSE) +
  tm_compass(type = "arrow",
             position = c(0.01, 0.1), 
             size = 1.5) +
  tm_scale_bar(size = 1, 
               position = c("left", "bottom")) )


tmap_save(tm = imperv.map, filename = paste0(Plots_wd, "/Map_study_area2.png"), width = NA, height = NA,
          units = NA, dpi = 600)
```
