---
title: "Run_bird_JSDM"
author: "Aimara Planillo"
date: "12/10/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Hierarchical Species Community Models for Bird abundance 

This scripts includes a bird data exploration and the code for running the 
bird abundance HSCM 


### Packages
```{r}
library(dplyr)
library(stringr)
library(Hmsc)
library(coda)
library(corrplot)
library(MCMCvis)
library(reshape2)
library(abind)
library(ggplot2)
library(sf)
library(tmap)
```

### Workspace
```{r}
WorkDir <- getwd()
output_wd <- paste0(WorkDir, "/5_Results/Bird_model/")
```


# Run Model Bird abundance

The main characteristics of the data used for the model are:
- 66 bird species
- 29 transects (mean values of the transect)
- 10 Covariates: 
    (noise.100m + pop.100m + tree.100m + open.green100m): *environmental covariates* 
    * prey.abu: *prey abundance*
- 7 Traits: body size (*log-transformed*) + diet (*4 categories*) + migrant (*3 categories*)
- Random: location of the transects

```{r prepare data for model}
# load bird data
birds_km <- read.csv("./4_processed_data/NEWBirds_abund_noexo_norare2_noaquatic_bykm.csv") # matrix with sites in rows and bird abund in columns
birds_km$site <- tolower(birds_km$site)

bird_data <- birds_km %>%
  select(-site)  # remove site
nrow(bird_data) # [1] 29 sites
ncol(bird_data) #[1] 66 species

# response matrix
Y.birds <- as.matrix(bird_data)

# Environmental data
env100m.tmp <- read.csv("./2_Raw_data/transects_allenvir_100m.csv")

# renaming covariates for the analysis
abiotic.covariates <- cbind.data.frame(site = env100m.tmp$routecode_2, 
                                       tree.100m = env100m.tmp$tree_cover,
                                       open.green100m = env100m.tmp$open_green,
                                       noise.100m = env100m.tmp$noise,
                                       pop.100m = env100m.tmp$pop_100m)
head(abiotic.covariates)
summary(abiotic.covariates)

# Prey data
carabids <- read.csv("./2_Raw_data/carabids_summary_pois_predictions.csv")
grasshoppers <- read.csv("./2_Raw_data/grasshoppers_summary_pois_predictions.csv")
spiders <- read.csv("./2_Raw_data/spiders_summary_pois_predictions.csv")

preys <- cbind.data.frame(site.preys = carabids$site,
                          carabids.abu = carabids$Abu,
                          grasshop.abu = grasshoppers$Abu,
                          spiders.abu = spiders$Abu,
                          carabids.rich = carabids$S,
                          grasshop.rich = grasshoppers$S,
                          spiders.rich = spiders$S)
preys

## Check correlations in predictors
# Correlations in Xdata
all.cov <- cbind.data.frame(abiotic.covariates, preys[,-1])
cor(all.cov[,2:ncol(all.cov)])

# All the abundances of prey are correlated, thus we use a summary of them 
all.cov$prey.abu <- all.cov$carabids.abu + all.cov$grasshop.abu + all.cov$spiders.abu

## Spatial coordinates
xyData.tmp <- read.csv('./2_Raw_data/transects_coordinates_25833.csv')
head(xyData.tmp)

xyData <- cbind(x = xyData.tmp$X, y = xyData.tmp$Y)

## Traits
# Load traits
Trdata.tmp <- read.csv("./4_Processed_data/NEWBirds_traits_noexo_norare2_noaquatic.csv") # matrix with species in rows and traits in columns
Trdata.tmp
nrow(Trdata.tmp) #[1] 66 species
head(Trdata.tmp)

# select only hte traits we are going to use and species names
Trdata <- Trdata.tmp %>% 
  select(scientific2, diet.5cat, migrant, bodymass.value)

# correct format of variables
Trdata$diet.5cat <- factor(Trdata$diet.5cat, 
                           levels = c("PlantSeed", "Invertebrate", "Omnivore", "VertFishScav"))

Trdata$migrant <- factor(Trdata$migrant, 
                           levels = c("Resident", "Semi", "Migrant" ))

Trdata$log10.bodymass <- log(Trdata$bodymass.value)


Trdata 
```


## Set up and Run the model


```{r}
### Study design
studyDesign <- data.frame(site = birds_km$site)
sRL <- xyData
rownames(sRL) <- birds_km$site
rL <- HmscRandomLevel(sData = sRL)
rL$s

### Define MCMC parameters
thin <- 50
samples <- 15000
transient <- 30000
nChains <- 3
verbose <- 10000

# Regression formula for environmental covariates
XFormula.mbirds = ~ (tree.100m + open.green100m + noise.100m + pop.100m) * prey.abu

# Regression formula for traits
TrFormula.mbirds = ~ migrant + diet.5cat + log10.bodymass


# set the model
## GAUSSIAN DISTRIBUTION: relative abundance data 
m <- Hmsc(Y = Y.birds, XData = all.cov, XFormula = XFormula.mbirds,
           studyDesign=studyDesign, ranLevels=list(site=rL),
          TrFormula = TrFormula.mbirds, TrData = Trdata,
          distr = "normal")

# run Markov Chains
m <- sampleMcmc(m, thin = thin, samples = samples, transient = transient,
                nChains = nChains, verbose = verbose)
#Save de model
# filename <- "model_birds_3sites_perkm.rds"
# saveRDS(m, file = paste0(output_wd, filename))
```


# Check Bird model

```{r model convergence}
# Evaluate convergence: Effective sample size and gelman-rubin diagnostic (potencial reduction factor)
mpost <- convertToCodaObject(m)
# Numerical output
ess.beta <- effectiveSize(mpost$Beta)
gd.beta <- gelman.diag(mpost$Beta, multivariate = FALSE)$psrf
ess.gamma <- effectiveSize(mpost$Gamma)
gd.gamma <- gelman.diag(mpost$Gamma, multivariate = FALSE)$psrf 
ess.omega <- effectiveSize(mpost$Omega[[1]])
gd.omega <- gelman.diag(mpost$Omega[[1]], multivariate = FALSE)$psrf
convergence.names <- c("ess.beta", "ess.gamma", "ess.omega",
               "gd.beta", "gd.gamma", "gd.omega")
convergence.list <- list(ess.beta, ess.gamma, ess.omega,
                 gd.beta, gd.gamma, gd.omega)
# save output
for (i in 1:length(convergence.names)){
  write.csv(convergence.list[[i]], paste0(output_wd, convergence.names[i], ".csv"))
}

# Graphical output
png("birds_model_convergence_hist.png", width = 800, height = 1000,
    pointsize = 20)
par(mfrow=c(3,2))
hist(ess.beta, main = "ess(beta)_birds")
hist(ess.gamma, main = "ess(gamma)_birds")
hist(ess.omega, main = "ess(omega)_birds")
hist(gd.beta, main = "psrf(beta)_birds")
hist(gd.gamma, main = "psrf(gamma)_birds")
hist(gd.omega, main = "psrf(omega)_birds")
dev.off()

# Save plots of the chains
MCMCtrace(mpost$Beta, 
          pdf = TRUE, 
          open_pdf = FALSE,
          filename = "birds_MCMCtrace_beta",
          wd = getwd())
MCMCtrace(mpost$Gamma, 
          pdf = TRUE, 
          open_pdf = FALSE,
          filename = "birds_MCMCtrace_gamma",
          wd = getwd())
MCMCtrace(mpost$Omega[[1]], 
          pdf = TRUE, 
          open_pdf = FALSE,
          filename = "birds_MCMCtrace_omega",
          wd = getwd())

par(mfrow=c(1,1))
```

```{r Model Fit R2}
# Explanatory R2. Get predictions for the observed values 
preds <- computePredictedValues(m, expected = TRUE)
preds.values <- apply(abind(preds,along=3),c(1,2), mean) # We use the mean because it was a normal model

# R2 with the built in function
modelr2.explanatory <- evaluateModelFit(hM = m, predY = preds)

# R2 of the model
mean(modelr2.explanatory$R2, na.rm = TRUE)

temp.r2 <- cbind.data.frame(r2 = as.numeric(modelr2.explanatory$R2), 
                            species = as.character(m$spNames))
plot(temp.r2$r2~temp.r2$species)

# R2 for species, Manually comparing observed vs predicted
R2.sp <- matrix(NA, m$ns, 1)
for (i in 1:m$ns) {
  R2.sp[i, ] <- cor(preds.values[, i],m$Y[, i])^2
}

mean(R2.sp, na.rm=TRUE)


# R2 for sites, Manually comparing observed vs predicted
R2.site <- matrix(NA, m$ny, 1)
for (i in 1:m$ny) {
  R2.site[i, ] <- cor(preds.values[i, ], m$Y[i, ])^2
}
mean(R2.site)
```


```{r estimates beta and gamma parameters}
# load premade function for plotting
source(paste0(WorkDir, "/3_Scripts/Function_PlotBetas.R"))

## Beta values
Beta.results <- as.data.frame(MCMCsummary(mpost$Beta))
postBeta <- getPostEstimate(m, parName = "Beta")

# default plots
plotBeta(m, post = postBeta, param = "Support", supportLevel = 0.95)
plotBeta(m, post = postBeta, param = "Support", supportLevel = 0.75)

# customized plots
# Print a plot for each predictor
plot.cov.betas(m, modelType = "birds") 


# plots of coefficients
C4 <- mpost$Beta[, grep("C4",colnames(mpost$Beta[[1]]))]
MCMCplot(C4, 
           ref_ovl = TRUE,
           rank = FALSE,
           xlab = 'noise',
           labels_sz = 0.3,
           med_sz = 1,
           thick_sz = 1,
           thin_sz = 1,
           ax_sz = 1,
           main_text_sz = 1)

## Plots of Coefficients of the Interactions
# extract the posteriors for the interactions
C7 <- mpost$Beta[, grep("C7",colnames(mpost$Beta[[1]]))]
colnames(C7[[1]])
head(C7)
C8 <- mpost$Beta[, grep("C8",colnames(mpost$Beta[[1]]))]
C9 <- mpost$Beta[, grep("C9",colnames(mpost$Beta[[1]]))]
C10 <- mpost$Beta[, grep("C10",colnames(mpost$Beta[[1]]))]

interactions <- list(C7, C8, C9, C10)
inter.names <- c("tree x prey", 
                 "open.green x prey",
                 "noise X prey",
                 "pop X prey")

# save the plots in a loop
for (i in 1:length(interactions)){
png(paste0("Betas_coef_plot_C", i+6, "_", inter.names[i], ".png"),
   width = 4000, height = 2500)
 MCMCplot(interactions[[i]], 
          ISB = FALSE,
          ref_ovl = TRUE,
          rank = FALSE,
          xlab = 'ESTIMATE',
          main = m$covNames[i+6],
          labels_sz = 2.5,
          med_sz = 4,
          thick_sz = 5,
          thin_sz = 3,
          ax_sz = 5,
          axis_text_sz = 3,
          tick_text_sz = 2, 
          main_text_sz = 4,
          mar = c(7, 4.1, 4.1, 2.1))
  dev.off()
}


## Gamma values
Gamma.results <- as.data.frame(MCMCsummary(mpost$Gamma))
# Default gamma plot in Hmsc package
postGamma <- getPostEstimate(m, parName = "Gamma")
plotGamma(hM = m, post = postGamma, param = "Support", supportLevel = 0.75)

# Coef. gammma
MCMCplot(mpost$Gamma, ref_ovl = TRUE)
```


```{r species co-occurrence}
# species associations in residual variance
OmegaCor <- computeAssociations(m)
OmegaCor[[1]]$mean
OmegaCor[[1]]$support

supportLevel <- 0.75

# Default plot in Hmsc package
toPlot <- ((OmegaCor[[1]]$support > supportLevel)
           + (OmegaCor[[1]]$support < (1 - supportLevel)) > 0) * OmegaCor[[1]]$mean
corrplot(toPlot, method = "color", 
         col = colorRampPalette(c("blue", "white", "red"))(200),
         title = paste0("random effect level: ", m$rLNames[1]), 
         mar = c(0,0,1,0))

par(mar = c (5,5,4,5))

# create data frame for customized plot
assoc.mean <- melt(OmegaCor[[1]]$mean)
assoc.support <- melt(OmegaCor[[1]]$support)

nrow(assoc.mean); nrow(assoc.support)
cbind.data.frame(assoc.mean$X1[c(1:50, 4150:4200)], assoc.support$X1[c(1:50, 4150:4200)])

associations <- cbind.data.frame(assoc.mean, assoc.support$value)
colnames(associations) <- c("species1", "species2", "mean", "support")

head(associations, 20)

write.csv(associations, paste0(output_wd, "Birds_associations.csv")
          , row.names = FALSE)
```


### Variance partitioning

The order of the variables, if they are continuous, is 
1. intercept(this can be in any group)
2. first variable
3. second variable
ETC.

The formulas we used for running the models area: 
XFormula.grasshop = ~ (tree.100m + open.green100m + noise.100m + pop.100m) * prey.abu

```{r variance partitioning}
# design matrix
head(m$X)

# Total variance explained by hte model can be partition into the contributions
# of each fixed effect (or group) and random effect
VP <- computeVariancePartitioning(m, group = c(1,1,1,2,2,3,4,4,4,4), groupnames = c("nature", "urban", "prey", "interactions"))

# Save the Variance partitioning info
VP$vals
VP$R2T

VP.table <- as.data.frame(VP$vals) 
VP.table
write.csv(VP.table, paste0(output_wd, "Varpart_values.csv"))

# plot var part. default plot
plotVariancePartitioning(m, VP = VP, las = 2, cex.names = 0.4)
title(main = "\n \nBirds")
```


# Bird model: Predict responses to environmental gradients

We are going to use 95CI
```{r working folder}
Gradients_wd <- paste0(WorkDir, "5_Results/Bird_EnvGradients/")
```


```{r create response gradients for species abundance}
# create a gradient of prey abundance
Gradient.prey <- constructGradient(m, focalVariable = "prey.abu")
predY.prey <- predict(m, XData = Gradient.prey$XDataNew, studyDesign = Gradient.prey$studyDesignNew, 
                 ranLevels = Gradient.prey$rLNew, expected = FALSE)
# saveRDS(predY.prey, paste0(Gradients_wd, "pred_response_prey.rds"))

# Summed response to prey.abu
plotGradient(m, Gradient.prey, pred=predY, measure="S", las=1,
             showData = TRUE, main='Species richness (measure="S")')
# Example one species: species 10
plotGradient(m, Gradient, pred=predY, measure="Y", index=10, las=1,
             showData = TRUE, main='Focal species occurrence (measure="Y")')

# Community-weighted mean values of traits
plotGradient(m, Gradient, pred=predY, measure="T", index=2, las=1,
             showData = TRUE, main='Mean trait value (measure="T")')
# save the plot
png(paste0(Gradients_wd, "/birds_totalabund_prey.png"))
plotGradient(m, Gradient.prey, pred=predY.prey, measure="S", las=1,
             showData = TRUE, main='Total abundance')
dev.off()

## we predict for all our variables
# Summed response to tree
Gradient.tree <- constructGradient(m, focalVariable = "tree.100m")
predY.tree <- predict(m, XData = Gradient.tree$XDataNew, 
                 studyDesign = Gradient.tree$studyDesignNew, 
                 ranLevels = Gradient.tree$rLNew, expected = FALSE)
# saveRDS(predY.tree, paste0(Gradients_wd, "pred_response_tree.rds"))
png(paste0(Gradients_wd, "/birds_totalabund_tree.png"))
plotGradient(m, Gradient.tree, pred=predY.tree, measure="S", las=1,
             showData = TRUE, main='Total abundance')
dev.off()

# Summed response to open.green
Gradient.og <- constructGradient(m, focalVariable = "open.green100m")
predY.og <- predict(m, XData = Gradient.og$XDataNew, 
                 studyDesign = Gradient.og$studyDesignNew, 
                 ranLevels = Gradient.og$rLNew, expected = FALSE)
# saveRDS(predY.og, paste0(Gradients_wd, "pred_response_ogreen.rds"))
png(paste0(Gradients_wd, "/birds_totalabund_ogreen.png"))
plotGradient(m, Gradient.og, pred=predY.og, measure="S", las=1,
             showData = TRUE, main='Total abundance')
dev.off()

# Summed response to noise
Gradient.noise <- constructGradient(m, focalVariable = "noise.100m")
predY.noise <- predict(m, XData = Gradient.noise$XDataNew, 
                 studyDesign = Gradient.noise$studyDesignNew, 
                 ranLevels = Gradient.noise$rLNew, expected = FALSE)
# saveRDS(predY.noise, paste0(Gradients_wd, "pred_response_noise.rds"))
png(paste0(Gradients_wd, "/birds_totalabund_noise.png"))
plotGradient(m, Gradient.noise, pred=predY.noise, measure="S", las=1,
             showData = TRUE, main='Total abundance')
dev.off()

# Summed response to human population density
Gradient.pop <- constructGradient(m, focalVariable = "pop.100m")
predY.pop <- predict(m, XData = Gradient.pop$XDataNew, 
                 studyDesign = Gradient.pop$studyDesignNew, 
                 ranLevels = Gradient.pop$rLNew, expected = FALSE)
# saveRDS(predY.pop, paste0(Gradients_wd, "pred_response_pop.rds"))
png(paste0(Gradients_wd, "/birds_totalabund_pop.png"))
plotGradient(m, Gradient.pop, pred=predY.pop, measure="S", las=1,
             showData = TRUE, main='Total abundance')
dev.off()
```


```{r plot response gradients for traits}
# save plots in a loop 

# Community-weigthed trait values responses to tree
for (i in 1:length(m$trNames)){
  png(paste0(Gradients_wd, "birds_trait_", m$trNames[i], "_tree.png"))
  plotGradient(m, Gradient.tree, pred=predY.tree, 
               measure="T", index = i, las=1,
               showData = TRUE, main=paste0('Trait response: ', m$trNames[i]))
  dev.off()
}

# Community-weigthed trait values response to open.green
for (i in 1:length(m$trNames)){
  png(paste0(Gradients_wd, "/birds_trait_", m$trNames[i], "_og.png"))
  plotGradient(m, Gradient.og, pred=predY.og, 
               measure="T", index = i, las=1,
               showData = TRUE, main=paste0('Trait response: ', m$trNames[i]))
  dev.off()
}

# Community-weigthed trait values response to noise
for (i in 1:length(m$trNames)){
  png(paste0(Gradients_wd, "/birds_trait_", m$trNames[i], "_noise.png"))
  plotGradient(m, Gradient.noise, pred=predY.noise, 
               measure="T", index = i, las=1,
               showData = TRUE, main=paste0('Trait response: ', m$trNames[i]))
  dev.off()
}

# Community-weigthed trait values response to human population density
for (i in 1:length(m$trNames)){
  png(paste0(Gradients_wd, "/birds_trait_", m$trNames[i], "_pop.png"))
  plotGradient(m, Gradient.pop, pred=predY.pop, 
               measure="T", index = i, las=1,
               showData = TRUE, main=paste0('Trait response: ', m$trNames[i]))
  dev.off()
}

# Community-weigthed trait values response to prey abundance
for (i in 1:length(m$trNames)){
  png(paste0(Gradients_wd, "/birds_trait_", m$trNames[i], "_prey.png"))
  plotGradient(m, Gradient.prey, pred=predY.prey, 
               measure="T", index = i, las=1,
               showData = TRUE, main=paste0('Trait response: ', m$trNames[i]))
  dev.off()
}
```


```{r response of species richness to environmental gradients}

## species richness response to tree cover
predY.treesp <- list()
for (i in 1:length(predY.tree)){
  predY.treesp[[i]] <- predY.tree[[i]]
  predY.treesp[[i]][predY.treesp[[i]] >0] <- 1
  predY.treesp[[i]][predY.treesp[[i]] <0] <- 0
}

png(paste0(Gradients_wd, "/birds_sprichness_tree.png"))
plotGradient(m, Gradient.tree, pred=predY.treesp, measure="S", las=1,
             showData = FALSE, main='Species richness')
dev.off()

# Sp richness response to open.green
predY.ogsp <- list()
for (i in 1:length(predY.og)){
  predY.ogsp[[i]] <- predY.og[[i]]
  predY.ogsp[[i]][predY.treesp[[i]] >0] <- 1
  predY.ogsp[[i]][predY.ogsp[[i]] <0] <- 0
}
png(paste0(Gradients_wd, "/birds_sprichness_ogreen.png"))
plotGradient(m, Gradient.og, pred=predY.ogsp, measure="S", las=1,
             showData = FALSE, main='Species richness')
dev.off()

# Sp richness response to noise
# predY.noise 
predY.noisesp <- list()
for (i in 1:length(predY.noise)){
  predY.noisesp[[i]] <- predY.noise[[i]]
  predY.noisesp[[i]][predY.noisesp[[i]] >0] <- 1
  predY.noisesp[[i]][predY.noisesp[[i]] <0] <- 0
}

png(paste0(Gradients_wd, "/birds_sprichness_noise.png"))
plotGradient(m, Gradient.noise, pred=predY.noisesp, measure="S", las=1,
             showData = FALSE, main='Species richness')
dev.off()

# Summed response to human population density
# predY.pop
predY.popsp <- list()
for (i in 1:length(predY.pop)){
  predY.popsp[[i]] <- predY.pop[[i]]
  predY.popsp[[i]][predY.popsp[[i]] >0] <- 1
  predY.popsp[[i]][predY.popsp[[i]] <0] <- 0
}
png(paste0(Gradients_wd, "/birds_sprichness_pop.png"))
plotGradient(m, Gradient.pop, pred=predY.popsp, measure="S", las=1,
             showData = FALSE, main='Species richness')
dev.off()

# Summed response to prey abundance
# predY.prey 
predY.preysp <- list()
for (i in 1:length(predY.prey)){
  predY.preysp[[i]] <- predY.prey[[i]]
  predY.preysp[[i]][predY.preysp[[i]] >0] <- 1
  predY.preysp[[i]][predY.preysp[[i]] <0] <- 0
}
png(paste0(Gradients_wd, "/birds_sprichness_prey.png"))
plotGradient(m, Gradient.prey, pred=predY.preysp, measure="S", las=1,
             showData = FALSE, main='Species richness')
dev.off()
```


## Responses to environmental variables at different values of prey abundance 

We are going to use three levels: 
low = 135 (minimun value of the index), 
mean = 200 (mean), 
high = 365 (maximum value of the index)

```{r select prey values}
lowprey <- min(m$XData$prey.abu)
meanprey <- mean(m$XData$prey.abu)
highprey <- max(m$XData$prey.abu)
prey.gradient <- cbind(lowprey, meanprey, highprey)
colnames(prey.gradient)
```


```{r response to noise and prey gradient}
for (i in 1:ncol(prey.gradient)){
  #construct the gradient for the predictions
  gradient <- constructGradient(m, focalVariable = "noise.100m",
                                        non.focalVariables = list(prey.abu=list(3,prey.gradient[,i])),
                                        ngrid = nrow(m$Y))
  gradname <- paste0("gradient.noise.", colnames(prey.gradient)[i])
  assign(gradname, gradient)
  # predict with the iterations
  predictions <- predict(m, XData = gradient$XDataNew, 
                           studyDesign = gradient$studyDesignNew,
                           ranLevels = gradient$rLNew, 
                           expected = FALSE)
  predname <- paste0("predY.noise.", colnames(prey.gradient)[i])
  assign(predname, predictions)
  # Get the final averaged predictions
  predS <- abind(lapply(predictions, rowSums), along = 2)
  predSname <- paste0("predS.noise.", colnames(prey.gradient)[i])
  assign(predSname, predS)
}

noise.gradient.list <- list(gradient.noise.lowprey, gradient.noise.meanprey, gradient.noise.highprey)
predY.noise.list <- list(predY.noise.lowprey, predY.noise.meanprey, predY.noise.highprey)
predS.noise.list <- list(predS.noise.lowprey, predS.noise.meanprey, predS.noise.highprey)


# saveRDS(noise.gradient.list, paste0(Gradients_wd, "/noise_gradient_list.rds"))
# saveRDS(predY.noise.list, paste0(Gradients_wd, "/noise_predY_gradient_list.rds"))
# saveRDS(predS.noise.list, paste0(Gradients_wd, "/noise_predCommunity_gradient_list.rds"))
```

```{r response to human population and prey gradient}
for (i in 1:ncol(prey.gradient)){
  gradient <- constructGradient(m, focalVariable = "pop.100m",
                                        non.focalVariables = list(prey.abu=list(3,prey.gradient[,i])),
                                        ngrid = nrow(m$Y))
  gradname <- paste0("gradient.pop.", colnames(prey.gradient)[i])
  assign(gradname, gradient)
  predictions <- predict(m, XData = gradient$XDataNew, 
                           studyDesign = gradient$studyDesignNew,
                           ranLevels = gradient$rLNew, 
                           expected = FALSE)
  predname <- paste0("predY.pop.", colnames(prey.gradient)[i])
  assign(predname, predictions)
  predS <- abind(lapply(predictions, rowSums), along = 2)
  predSname <- paste0("predS.pop.", colnames(prey.gradient)[i])
  assign(predSname, predS)
}

pop.gradient.list <- list(gradient.pop.lowprey, gradient.pop.meanprey, gradient.pop.highprey)
predY.pop.list <- list(predY.pop.lowprey, predY.pop.meanprey, predY.pop.highprey)
predS.pop.list <- list(predS.pop.lowprey, predS.pop.meanprey, predS.pop.highprey)
```

```{r response to tree cover and prey gradient}
for (i in 1:ncol(prey.gradient)){
  gradient <- constructGradient(m, focalVariable = "tree.100m",
                                        non.focalVariables = list(prey.abu=list(3,prey.gradient[,i])),
                                        ngrid = nrow(m$Y))
  gradname <- paste0("gradient.tree.", colnames(prey.gradient)[i])
  assign(gradname, gradient)
  predictions <- predict(m, XData = gradient$XDataNew, 
                           studyDesign = gradient$studyDesignNew,
                           ranLevels = gradient$rLNew, 
                           expected = FALSE)
  predname <- paste0("predY.tree.", colnames(prey.gradient)[i])
  assign(predname, predictions)
  predS <- abind(lapply(predictions, rowSums), along = 2)
  predSname <- paste0("predS.tree.", colnames(prey.gradient)[i])
  assign(predSname, predS)
}

tree.gradient.list <- list(gradient.tree.lowprey, gradient.tree.meanprey, gradient.tree.highprey)
predY.tree.list <- list(predY.tree.lowprey, predY.tree.meanprey, predY.tree.highprey)
predS.tree.list <- list(predS.tree.lowprey, predS.tree.meanprey, predS.tree.highprey)
```

```{r response to open green and prey gradient}
# Make the predictions for the three levels of prey abund
for (i in 1:ncol(prey.gradient)){
    #construct the gradient for the predictions
  gradient <- constructGradient(m, focalVariable = "open.green100m",
                                        non.focalVariables = list(prey.abu=list(3,prey.gradient[,i])),
                                        ngrid = nrow(m$Y))
  gradname <- paste0("gradient.open.green.", colnames(prey.gradient)[i])
  assign(gradname, gradient)
  # predict with the iterations
  predictions <- predict(m, XData = gradient$XDataNew, 
                           studyDesign = gradient$studyDesignNew,
                           ranLevels = gradient$rLNew, 
                           expected = FALSE)
  predname <- paste0("predY.open.green.", colnames(prey.gradient)[i])
  assign(predname, predictions)
  # Get the final averaged predictions
  predS <- abind(lapply(predictions, rowSums), along = 2)
  predSname <- paste0("predS.open.green.", colnames(prey.gradient)[i])
  assign(predSname, predS)
}

ogreen.gradient.list <- list(gradient.open.green.lowprey, gradient.open.green.meanprey, gradient.open.green.highprey)
predY.ogreen.list <- list(predY.open.green.lowprey, predY.open.green.meanprey, predY.open.green.highprey)
predS.ogreen.list <- list(predS.open.green.lowprey, predS.open.green.meanprey, predS.open.green.highprey)
```


## Create hte data frames for plotting


```{r extract values for plotting noise and prey}
prey.gradient.names <- c("lowprey", "meanprey", "highprey")

# extract env gradient
noise.gradient <- gradient.noisexprey[[1]]$XDataNew$noise.100m

# Get Credible intervals for the three community predicitons
for (i in 1:length(predS.noisexprey)){
    qpred =  apply(predS.noisexprey[[i]], c(1), quantile, prob = c(0.025, 0.125, 0.5, 0.875, 0.975), na.rm = TRUE)
    assign(paste0("qpred.noise.", prey.gradient.names[i]), qpred)
      }

# Get Y X values from the model, to plot the points
pY = rowSums(m$Y)
pX = m$XData[, "noise.100m"]

# Build dataframes with all the data
toplot.noise.lowprey <- cbind.data.frame(prey = gradient.noisexprey[[1]]$XDataNew$prey.abu,
                                         noise = noise.gradient,
                                         noise.obs = pX,
                                         sp.obs = pY,
                                         beta = qpred.noise.lowprey[3,],
                                         low95 = qpred.noise.lowprey[1,], 
                                         low75 = qpred.noise.lowprey[2,], 
                                         high75 = qpred.noise.lowprey[4,], 
                                         high95 = qpred.noise.lowprey[5,])

toplot.noise.meanprey <- cbind.data.frame(prey = gradient.noisexprey[[2]]$XDataNew$prey.abu,
                                         noise = noise.gradient,
                                         noise.obs = pX,
                                         sp.obs = pY,
                                         beta = qpred.noise.meanprey[3,],
                                         low95 = qpred.noise.meanprey[1,], 
                                         low75 = qpred.noise.meanprey[2,], 
                                         high75 = qpred.noise.meanprey[4,], 
                                         high95 = qpred.noise.meanprey[5,])

toplot.noise.highprey <- cbind.data.frame(prey = gradient.noisexprey[[3]]$XDataNew$prey.abu,
                                         noise = noise.gradient,
                                         noise.obs = pX,
                                         sp.obs = pY,
                                         beta = qpred.noise.highprey[3,],
                                         low95 = qpred.noise.highprey[1,], 
                                         low75 = qpred.noise.highprey[2,], 
                                         high75 = qpred.noise.highprey[4,], 
                                         high95 = qpred.noise.highprey[5,])

toplot.noise.all <- rbind(toplot.noise.lowprey, toplot.noise.meanprey, toplot.noise.highprey)
toplot.noise.all$site <- rownames(toplot.noise.all)

write.csv(toplot.noise.all, paste0(Gradients_wd, "/Toplot_noisexprey_sites.csv"), row.names = FALSE)
```

```{r extract values for plotting human population and prey}
# extract env gradient
pop.gradient <- gradient.popxprey[[1]]$XDataNew$pop.100m

# Get Credible intervals for the three community predicitons
for (i in 1:length(predS.popxprey)){
    qpred =  apply(predS.popxprey[[i]], c(1), quantile, prob = c(0.025, 0.125, 0.5, 0.875, 0.975), na.rm = TRUE)
    assign(paste0("qpred.pop.", prey.gradient.names[i]), qpred)
      }
  
qpred.pop.lowprey  

# Get Y X values from the model, to plot the points
pY = rowSums(m$Y)
pX = m$XData[, "pop.100m"]

# Build dataframes with all the data
toplot.pop.lowprey <- cbind.data.frame(prey = gradient.popxprey[[1]]$XDataNew$prey.abu,
                                         pop = pop.gradient,
                                         pop.obs = pX,
                                         sp.obs = pY,
                                         beta = qpred.pop.lowprey[3,],
                                         low95 = qpred.pop.lowprey[1,], 
                                         low75 = qpred.pop.lowprey[2,], 
                                         high75 = qpred.pop.lowprey[4,], 
                                         high95 = qpred.pop.lowprey[5,])

toplot.pop.meanprey <- cbind.data.frame(prey = gradient.popxprey[[2]]$XDataNew$prey.abu,
                                         pop = pop.gradient,
                                         pop.obs = pX,
                                         sp.obs = pY,
                                         beta = qpred.pop.meanprey[3,],
                                         low95 = qpred.pop.meanprey[1,], 
                                         low75 = qpred.pop.meanprey[2,], 
                                         high75 = qpred.pop.meanprey[4,], 
                                         high95 = qpred.pop.meanprey[5,])

toplot.pop.highprey <- cbind.data.frame(prey = gradient.popxprey[[3]]$XDataNew$prey.abu,
                                         pop = pop.gradient,
                                         pop.obs = pX,
                                         sp.obs = pY,
                                         beta = qpred.pop.highprey[3,],
                                         low95 = qpred.pop.highprey[1,], 
                                         low75 = qpred.pop.highprey[2,], 
                                         high75 = qpred.pop.highprey[4,], 
                                         high95 = qpred.pop.highprey[5,])

toplot.pop.all <- rbind(toplot.pop.lowprey, toplot.pop.meanprey, toplot.pop.highprey)
toplot.pop.all$site <- rownames(toplot.pop.all)

write.csv(toplot.pop.all, paste0(Gradients_wd, "/Toplot_popxprey_sites.csv"), row.names = FALSE)
```

```{r extract values for plotting tree cover and prey}
# extract env gradient
tree.gradient <- gradient.treexprey[[1]]$XDataNew$tree.100m

# Get Credible intervals for the three community predicitons
for (i in 1:length(predS.treexprey)){
    qpred =  apply(predS.treexprey[[i]], c(1), quantile, prob = c(0.025, 0.125, 0.5, 0.875, 0.975), na.rm = TRUE)
    assign(paste0("qpred.tree.", prey.gradient.names[i]), qpred)
      }
  
qpred.tree.lowprey  

# Get Y X values from the model, to plot the points
pY = rowSums(m$Y)
pX = m$XData[, "tree.100m"]

# Build dataframes with all the data
toplot.tree.lowprey <- cbind.data.frame(prey = gradient.treexprey[[1]]$XDataNew$prey.abu,
                                         tree = tree.gradient,
                                         tree.obs = pX,
                                         sp.obs = pY,
                                         beta = qpred.tree.lowprey[3,],
                                         low95 = qpred.tree.lowprey[1,], 
                                         low75 = qpred.tree.lowprey[2,], 
                                         high75 = qpred.tree.lowprey[4,], 
                                         high95 = qpred.tree.lowprey[5,])

toplot.tree.meanprey <- cbind.data.frame(prey = gradient.treexprey[[2]]$XDataNew$prey.abu,
                                         tree = tree.gradient,
                                         tree.obs = pX,
                                         sp.obs = pY,
                                         beta = qpred.tree.meanprey[3,],
                                         low95 = qpred.tree.meanprey[1,], 
                                         low75 = qpred.tree.meanprey[2,], 
                                         high75 = qpred.tree.meanprey[4,], 
                                         high95 = qpred.tree.meanprey[5,])

toplot.tree.highprey <- cbind.data.frame(prey = gradient.treexprey[[3]]$XDataNew$prey.abu,
                                         tree = tree.gradient,
                                         tree.obs = pX,
                                         sp.obs = pY,
                                         beta = qpred.tree.highprey[3,],
                                         low95 = qpred.tree.highprey[1,], 
                                         low75 = qpred.tree.highprey[2,], 
                                         high75 = qpred.tree.highprey[4,], 
                                         high95 = qpred.tree.highprey[5,])

toplot.tree.all <- rbind(toplot.tree.lowprey, toplot.tree.meanprey, toplot.tree.highprey)
toplot.tree.all$site <- rownames(toplot.tree.all)

write.csv(toplot.tree.all, paste0(Gradients_wd, "/Toplot_treexprey_sites.csv"), 
          row.names = FALSE)
```

```{r extract values for plotting open green area and prey}
# extract env gradient
ogreen.gradient <- gradient.ogreenxprey[[1]]$XDataNew$open.green100m

# Get Credible intervals for the three community predicitons
for (i in 1:length(predS.ogreenxprey)){
    qpred =  apply(predS.ogreenxprey[[i]], c(1), quantile, prob = c(0.025, 0.125, 0.5, 0.875, 0.975), na.rm = TRUE)
    assign(paste0("qpred.ogreen.", prey.gradient.names[i]), qpred)
      }
  
qpred.ogreen.lowprey  

# Get Y X values from the model, to plot the points
pY = rowSums(m$Y)
pX = m$XData[, "open.green100m"]

# Build dataframes with all the data
toplot.ogreen.lowprey <- cbind.data.frame(prey = gradient.ogreenxprey[[1]]$XDataNew$prey.abu,
                                         ogreen = ogreen.gradient,
                                         ogreen.obs = pX,
                                         sp.obs = pY,
                                         beta = qpred.ogreen.lowprey[3,],
                                         low95 = qpred.ogreen.lowprey[1,], 
                                         low75 = qpred.ogreen.lowprey[2,], 
                                         high75 = qpred.ogreen.lowprey[4,], 
                                         high95 = qpred.ogreen.lowprey[5,])

toplot.ogreen.meanprey <- cbind.data.frame(prey = gradient.ogreenxprey[[2]]$XDataNew$prey.abu,
                                         ogreen = ogreen.gradient,
                                         ogreen.obs = pX,
                                         sp.obs = pY,
                                         beta = qpred.ogreen.meanprey[3,],
                                         low95 = qpred.ogreen.meanprey[1,], 
                                         low75 = qpred.ogreen.meanprey[2,], 
                                         high75 = qpred.ogreen.meanprey[4,], 
                                         high95 = qpred.ogreen.meanprey[5,])

toplot.ogreen.highprey <- cbind.data.frame(prey = gradient.ogreenxprey[[3]]$XDataNew$prey.abu,
                                         ogreen = ogreen.gradient,
                                         ogreen.obs = pX,
                                         sp.obs = pY,
                                         beta = qpred.ogreen.highprey[3,],
                                         low95 = qpred.ogreen.highprey[1,], 
                                         low75 = qpred.ogreen.highprey[2,], 
                                         high75 = qpred.ogreen.highprey[4,], 
                                         high95 = qpred.ogreen.highprey[5,])

toplot.ogreen.all <- rbind(toplot.ogreen.lowprey, toplot.ogreen.meanprey, toplot.ogreen.highprey)
toplot.ogreen.all$site <- rownames(toplot.ogreen.all)

write.csv(toplot.ogreen.all, paste0(Gradients_wd, "/Toplot_ogreenxprey_sites.csv"),
          row.names = FALSE)
```
